Response to Reviewers

### 2025-07-09 paper/kobakian-cook-anzjs.tex
still to finish conclusion and working on the reason summary

\usepackage{pmboxdraw}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics},
  pdfauthor={Stephanie Kobakian; Dianne Cook},
  pdfkeywords={statistics, visual inference, geospatial, population},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\title{Comparing the Effectiveness of the Choropleth Map with a Hexagon
Tile Map for Communicating Cancer Statistics}

@@

\keywords{statistics; visual inference; geospatial; population}

@@

The experiment was conducted using the lineup protocol, a visual
inference procedure \citep{BCHLLSW09, GIIV}, that can be used to
objectively test the effectiveness of the two displays \citep{GTPCCD}. A
lineup embeds the data plot among a field of null data plots, and an
independent observer is asked to select the most different plot. It was
shown by \citet{VVSIALM} to be an effective way to conduct a hypothesis
test where a plot is treated as a test statistic. If the data plot is
selected it is analogous to a rejection of the null hypothesis (no
structure), and the observed data plot is unlikely to arise from the
null scenario.

@@

A cartogram is a general solution for adequately displaying a
population-based statistic. It transforms the geographic map base to
reflect the population in the geographic region, while preserving some
aspects of the geographic location. There are several cartogram
algorithms \citep{ACTUC, CBATCC}; each involves shifting the boundaries
of geographic units, using the value of the statistic to increase or
decrease the area taken by the geographic unit on the map. The changes
to the boundaries result in cartograms that accurately communicate
population by map area for each of the geographic units but can result
in losing the familiar geographic information. For Australia, the
transformations warp the country so that it is no longer recognizable.

@@

The hexagon tile map algorithm, automatically matches spatial regions to
their nearest hexagon tile, from a grid of tiles. It has the effect of
spreading out the inner city areas while maintaining the spatial
locations or regions in remote areas. The algorithm is available in the
R package, sugarbag \citep{sugarbag}. Figure~\ref{fig-thyroid} b shows
the hexagon tile map, where the map is coloured from substantially below
average (blue) to substantially above average (red) rates. The inner
city areas have expanded, making it possible to see the cancer incidence
in the small, densely populated areas. Remote regions are represented by
isolated hexagons, which is not ideal, but maintains the spatial
location of these data values. It is of interest to know how well the
spatial distribution patterns are seen from this display, in comparison
to how they are seen from the choropleth.

@@

In order to assess the effectiveness of the hexagon tile map, the lineup
protocol \citep{BCHLLSW09, GIIV} from visual inference procedures is
employed. The approach mirrors classical statistical inference. The
procedures for doing a power comparison of competing plot designed,
outlined in \citet{GTPCCD}, are followed.

@@

@@

\begin{itemize}
\tightlist
\item
  Map type: \emph{choropleth, hexagon tile}
\item
  Trend: \emph{South-East to North-West; Locations in three population
  centres; Locations in multiple population centres, }
\end{itemize}

@@

The R package \texttt{gstat} \citep{gstat} was used to simulate 144 null
sets, 12 data sets for each plot in a lineup, and 12 sets for 12
lineups. Each null set was further processed to produce maps with
similar spatial dependence as those produced by the methods in
\citet{atlas-methods}, by averaging a small number of spatial
neighbours.

@@

One of the lineup locations was chosen to embed the real trend model
map, in each of the four replicates, for the three trend models. The
locations were chosen by random sampling. Using random locations reduces
the chance that participants might inducing the location coincidentally.
Locations 1, 7, 10 and 11 were not in the sample.

The lineup locations were the same for both map types, because each set
of lineup data was used to produce a choropleth map lineup and hexagon
tile map lineup. Lineups were grouped into A or B, so that a participant
saw only one version. Participants were assigned to group A or B,
randomly, and thus evaluated either the choropleth or the hexagon tile
map lineup. Because there were four replicates of each lineup, each
participant evaluated two choropleth and two hexagon tile map lineups,
for each trend model. This design is illustrated in
Figure~\ref{fig-exp-design}.

@@

Each participant used the internet to access the survey. The data
transfer from the web application to the data set took place using a
secure link to the googlesheet used to store results. The application
connected to the googlesheet using the googlesheets \citep{sheets} R
package when participants opened the application, and interacted again
when participants chose to submit the survey. At this time it added the
participant's responses to the twelve lineup displays as twelve rows of
data in the googlesheet.

@@

Participants were recruited from the Figure Eight crowdsourcing platform
\citep{figeight} to evaluate lineups. The lineup protocol expects that
the participants are uninvolved judges with no prior knowledge of the
data, to avoid inadvertently affecting results. Potential participants
needed to have achieved level 2 or level 3 from prior work on the
platform. All participants were at least 18 years old.

@@

Basic descriptive statistics were used to contrast the detection rate
for the two types of displays. Comparison was also made across the trend
models, contrasting the mean and standard detection rate for each group,
who had seen the different map display type for each replicate.

Side-by-side dot plots were made of accuracy (efficiency) against map
type, faceted by trend model type.

Similar plots were made of the feedback and demographic variables -
reason for choice, reported difficulty, gender, age, education, having
lived in Australia - against the design variables.

@@

Of the 90 participants, 66 were male, and 24 female. Most participants
(55) had a Bachelors degree, 13 had a Masters degree, and the remaining
22 had high school diplomas. The age distribution was 13, 36, 21, 11, 6
for age groups 18-24, 25-34, 35-44, 45-54, 55+, with 3 preferring not to
answer. Only 1 reported having lived in Australia.

@@

Table~\ref{tbl-desc-stats} shows the means and standard deviations of
the detection rate for each type of plot and each trend model. This also
gives the standard deviations, the smallest standard deviation for all
sets of replicates was the Three Cities trend model shown in a
Choropleth display. This group of displays had a very small detection
rate of 0.04. The mean detection rate for the Three Cities trend model
shown as choropleth map lineups was also the smallest at 0.40. The
North-West to South-East (NW-SE) trend model unexpectedly had a higher
mean detection rate for the hexagon tile map displays, but the
difference in the means of detection rate was only 0.10.

\caption{\label{tbl-desc-stats}The mean and standard deviation of the
rate of detection for each trend model, calculated for the choropleth
and hexagon tile map displays.}

\begin{tabular}{lccc}
\toprule
Type & NW-SE & Three Cities & All Cities\\
\midrule
Choro. & 0.52 & 0.04 & 0.23\\
 & (0.50) & (0.19) & (0.42)\\
\addlinespace
Hex. & 0.62 & 0.40 & 0.58\\
 & (0.49) & (0.49) & (0.49)\\
\bottomrule
\end{tabular}

Table~\ref{tbl-detect-glmer1} presents a summary of the generalised
linear mixed effects model, testing the effect of map type and trend
model on the detection rate. The results support the summary from
Figure~\ref{fig-detect-compare} and all parameters are statistically
significant despite the large standard deviations observed in
Table~\ref{tbl-desc-stats}. Overall, the hexagon tile map performs
marginally better than the choropleth for all trend models, which is a
pleasant surprise. Allowing for the interaction effect, the difference
in detection rate decreases for population related displays for a
choropleth map lineup, but increases for a hexagon tile map display. The
log odds of detection show in Table~\ref{tbl-detect-glmer1} can be back
transformed after taking the sum of all terms for the trend and type of
display that are of interest. For the NW-SE distribution, the predicted
detection rate for the hexagon tile map display increases the predicted
probability of detection to 0.63 from 0.52 for choropleths, this is
almost exactly the difference seen in the table of means and is
significant only at the 0.05 level.

When a choropleth map display is used, the predicted detection rate for
the Three Cities trend, 0.03 is extremely low, especially compared to
the NW-SE trend of 0.52. When the All Cities trend is presented in a
choropleth display the predicted probability of detection is 0.22. The
hexagon tile map has a substantially high detection rate for the display
of a Three Cities trend 0.39 and All Cities trend 0.59.

@@

@@

\caption{\label{tbl-reason}The main reason that participants wrote for
their choice of plot when looking at each trend model shown in
choropleth and hexagon tile maps.}

@@

\begin{tabular}{llll}
\toprule
Trend & Detected & Choro. & Hex.\\
\midrule
 & No & trend & clusters\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash All Cities} & Yes & clusters, consistent & clusters\\
\cmidrule{1-4}
 & No & trend & clusters\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash Three Cities} & Yes & consistent & clusters\\
\cmidrule{1-4}
 & No & trend & clusters\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash NW-SE} & Yes & trend & clusters\\
\bottomrule
\end{tabular}

@@

@@

The differences seen in the Figure~\ref{fig-detect-compare} plot and
Table~\ref{tbl-desc-stats} are reflected in the model results.
Surprisingly the difference for the geographic distribution was
significant at the 0.05 level. It also showed that the hexagon tile map
display performs marginally better than the choropleth for all trend
models. Unexpectedly the detection rate suffers when using a choropleth
map to display population related distributions.

@@

Add Election results data + plus other statistics, other hexagon layout
software

>>>>

\usepackage{pmboxdraw}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Patterns in Australian Spatial Statistics},
  pdfauthor={Stephanie Kobakian; Dianne Cook},
  pdfkeywords={data visualisation, visual
inference, geospatial, statistical graphics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\title{Comparing the Effectiveness of the Choropleth Map with a Hexagon
Tile Map for Communicating Patterns in Australian Spatial Statistics}

@@

\keywords{data visualisation; visual inference; geospatial; statistical
graphics}

@@

The experiment was conducted using the lineup protocol, a visual
inference procedure \citep{BCHLLSW09, GIIV}, that can be used to
objectively test the effectiveness of the two displays \citep{GTPCCD}. A
lineup embeds the data plot among a field of null data plots, and an
independent observer is asked to select the most different plot. It was
shown by \citet{VVSIALM} to be an effective way to conduct a hypothesis
test where a plot is treated as a test statistic, and utilised for this
purpose in numerous studies \citep{FFS, Green2021, LCTV}. If the data
plot is selected it is analogous to a rejection of the null hypothesis
(specifying no structure), and the observed data plot is unlikely to
arise from the null scenario. The work of \citet{VH2016} compares the
lineup protocol to performance on statndard tests of visual ability,
concluding that participants' performance is related to general visual
aptitude, to classification rather than spatial reasoning.

@@

A cartogram is a general solution for adequately displaying a
population-based statistic. It transforms the geographic map base to
reflect the population in the geographic region, while preserving some
aspects of the geographic location. There are several cartogram
algorithms \citep{ACTUC, CBATCC}; each involves shifting the boundaries
of geographic units, using the value of the statistic to increase or
decrease the area taken by the geographic unit on the map. The changes
to the boundaries result in cartograms that accurately communicate
population by map area for each of the geographic units but can result
in losing the familiar geographic information. For Australia, the
transformations warp the country so that it is no longer recognisable
(see \citet{KCR} for details).

@@

The hexagon tile map algorithm, automatically matches spatial regions to
their nearest hexagon tile, from a grid of tiles. It has the effect of
spreading out the inner city areas while maintaining the spatial
locations or regions in remote areas. The algorithm is available in the
R package, sugarbag \citep{sugarbag}. Figure~\ref{fig-thyroid} b shows
the hexagon tile map, where the map is coloured from low (blue) to high
(red). The inner city areas have expanded, making it possible to see the
cancer incidence in the small, densely populated areas. Remote regions
are represented by isolated hexagons, which is not ideal, but maintains
the spatial location of these data values. It is of interest to know how
well the spatial distribution patterns are seen from this display, in
comparison to how they are seen from the choropleth.

@@

In order to assess the effectiveness of the hexagon tile map, the lineup
protocol \citep{BCHLLSW09, GIIV} from visual inference procedures is
employed. The approach mirrors classical statistical inference. The
procedures for doing a power comparison of competing plot designed,
outlined in \citet{GTPCCD}, are followed. It is the only current human
subjects testing protocol which quantitatively compares plot designs on
the basis of detection of structure relative to null distributions
\citep{VCH, VanderPlas2021Designing}. The premise for comparing two
designs is that the only difference between the two lineups is the plot
design, and hence difference in detection rate and time to detect is due
to the effectiveness of the plot design for differentiating between data
plot and null plots. The protocol has been used in to quantitatively
test plot designs numerous studies
\citep[e.g.][]{LHC, VH2016, KTV, RS2021}.

@@

There are several practical considerations when deploying the lineup
protocol: (1) determining the appropriate null distribution to compute
null samples, (2) employing independent observers to conduct the
evaluations, (3) varying location of the data plot in a lineup, (4) how
many null sets to include in the lineup, (5) construction of all lineups
in the experiment (see \citet{VRCH}), (6) questions presented to
participants to solicit evaluations, in addition to the usual
experimental design issues. These are described in detail in the next
section.

@@

\begin{itemize}
\item
  Map type: \emph{Choropleth, Hexagon tile}
\item
  Trend:

  \begin{itemize}
  \tightlist
  \item
    \emph{NW-SE}: Large spatial trend running diagonally across
    Australia
  \item
    \emph{Three Cities}: Locations in three population centres
  \item
    \emph{All Cities}: Locations in all state and territory capitals
  \end{itemize}
\end{itemize}

@@

The R package \texttt{gstat} \citep{gstat} was used to simulate 144 null
sets, 12 data sets for each plot in a lineup, and 12 sets for 12
lineups. Simulating spatial dependence is difficult as discussed by
\citet{BDMSTW} as a blockage for using the lineup protocol for testing
map displays. Because the result from \texttt{gstat} was inadequate to
mirror the spatial dependence patterns results in the \citet{atlas},
each null set was further smoothed. This was done by averaging a small
number of spatial neighbours, approximating the methods described in
\citet{atlas-methods}.

@@

One of the plot locations (1-20) is chosen to embed the data plot, in
each of the four replicates, for the three trend models. These locations
were chosen by random sampling. Using random locations reduces the
chance that participants might inducing the location coincidentally.
Locations 1, 7, 10 and 11 were not in the sample. \citet{YMCCSG} used
the lineup protocol for a genomics study and similarly varied the the
location of the data plot in replicates of treatments. Their results
demonstrated that the actual location of the data plot didn't affect
performance. So we don't expect that the location affects results but
randomising locations among different lineups is to guard against
participants expecting the data plot to be in a particular location.

The lineup locations were the same for both map types, because each set
of lineup data was used to produce a choropleth map lineup and hexagon
tile map lineup. This ensures that performance on the two map types can
be directly compared. Lineups were grouped into A or B, so that a
participant saw only one version. Participants were assigned to group A
or B, randomly, and thus evaluated either the choropleth or the hexagon
tile map lineup. Because there were four replicates of each lineup, each
participant evaluated two choropleth and two hexagon tile map lineups,
for each trend model. This design is illustrated in
Figure~\ref{fig-exp-design}.

@@

Each participant used the internet to access the survey, and data was
transferred by secure link from the web app to a Google sheet using the
\texttt{googlesheets} package tools \citep{sheets}.

@@

Participants were recruited from the Figure Eight crowdsourcing platform
\citep{figeight} to evaluate lineups. The lineup protocol expects that
the participants are uninvolved judges with no prior knowledge of the
data, to avoid inadvertently affecting results. Potential participants
needed to have achieved level 2 or level 3 from prior work on the
platform, ensuring only participants with a good record on prior tasks
could provide evaluations. All participants were at least 18 years old.

@@

Basic descriptive statistics were computed for the different
experimental treatments. Basic plots summarising detection rates by map
type and trend model type, and feedback and demographic variables
against the different experimental design elements.

@@

@@

@@

\caption{\label{tbl-detect-prop}Model estimates for the proportion of
detection in each of the trend models (standard error). Note that
selecting the data plot by chance would produce a detection rate of
0.05.}

\begin{tabular}{lccc}
\toprule
Map type & All Cities & Three Cities & NW-SE\\
\midrule
Choro. & 0.22 & 0.03 & 0.52\\
 & (0.03) & (0.01) & (0.04)\\
\addlinespace
Hex. & 0.59 & 0.39 & 0.63\\
 & (0.04) & (0.04) & (0.04)\\
\bottomrule
\end{tabular}

Table~\ref{tbl-detect-glmer1} presents a summary of the generalised
linear mixed effects model, testing the effect of map type and trend
model on the detection rate. The results support the summary from
Figure~\ref{fig-detect-compare}. Overall, the hexagon tile map performs
marginally better than the choropleth for all trend models, and with
differing magnitudes of effects. The subject-specific random intercepts
have mean 0.3 and standard devation 0.55.

The estimated detection proportion from the model fit, computed using
the \texttt{emmeans} package \citep{Lenth2025} are shown in
Table~\ref{tbl-detect-prop}. For the All Cities trend participants were
about three times more likely to detect the cluster pattern with the
hexagon tile map than the choropleth. The Three Cities trend was
practically not detectable in the choropleth map. The dection rates were
more similar for the NW-SE trend, but slightly higher for the hexagon
tile map, which was a surprise. Note that, these detection rates are all
substantially higher than chance, except for the choropleth map on the
Three Cities. For a single evaluation the detection rate of the data
plot selected by chance is \(1/20=0.05\) because the lineups used in
this experiment had 20 plots. When designing an experiment like this it
is important to produce lineups that strike a balance between simple and
hard, so that there is a chance of discovering the effect of interest.
This experiment has managed to do this extremely well, which is a result
of pilot studies, careful null data generation, and power calculations
to determine appropriate sample size.

@@

\caption{\label{tbl-reason}Summary of the reasons that participants
wrote for their choice of plot when looking at each trend model shown in
choropleth and hexagon tile maps.}

@@

\begin{tabular}{lllrrrrr}
\toprule
Trend & Detected & type & clusters & consistent & hotspots & no reason & trend\\
\midrule
 & No & Choro. & 50 & 12 & 11 & 9 & 56\\
\cmidrule{2-8}
 & No & Hex. & 27 & 6 & 5 & 15 & 22\\
\cmidrule{2-8}
 & Yes & Choro. & 14 & 14 & 3 & 4 & 7\\
\cmidrule{2-8}
\multirow{-4}{*}{\raggedright\arraybackslash All Cities} & Yes & Hex. & 44 & 4 & 14 & 8 & 35\\
\cmidrule{1-8}
 & No & Choro. & 54 & 12 & 17 & 16 & 74\\
\cmidrule{2-8}
 & No & Hex. & 42 & 6 & 11 & 15 & 34\\
\cmidrule{2-8}
 & Yes & Choro. & 1 & 4 & 0 & 0 & 2\\
\cmidrule{2-8}
\multirow{-4}{*}{\raggedright\arraybackslash Three Cities} & Yes & Hex. & 35 & 0 & 10 & 4 & 23\\
\cmidrule{1-8}
 & No & Choro. & 20 & 14 & 11 & 8 & 34\\
\cmidrule{2-8}
 & No & Hex. & 27 & 3 & 5 & 15 & 18\\
\cmidrule{2-8}
 & Yes & Choro. & 32 & 2 & 15 & 7 & 37\\
\cmidrule{2-8}
\multirow{-4}{*}{\raggedright\arraybackslash NW-SE} & Yes & Hex. & 56 & 12 & 6 & 6 & 32\\
\bottomrule
\end{tabular}

@@

Of the 90 participants, 66 were male, and 24 female. Most participants
(55) had a Bachelors degree, 13 had a Masters degree, and the remaining
22 had high school diplomas. The age distribution was 13, 36, 21, 11, 6
for age groups 18-24, 25-34, 35-44, 45-54, 55+, with 3 preferring not to
answer. Only 1 reported having lived in Australia. Note that, the
purpose of reporting these numbers is to illustrate the reasonable
variety of demographic background of participants. However, \$n=\$90 is
insufficient fit a model including this demographic information.
\citet{MHC} (published in 2025 but conducted in 2012) showed that there
is little difference in performance on lineup experiments between
demographic groups. \citet{VH2016} found that visual aptitude for
reading data plots was associated with math skills, but this study was
done on statistical plots of data, not maps. Assessing math ability
requires substantially more data collection, and does not necessarily
correlate with education level.

Summary statistics (not included here, but available in the analysis
code) show no differences in results between sets A and B in detection
rate or time taken. Similarly, detection rates and time taken vary
little across age, education and gender.

@@

The differences seen in the Figure~\ref{fig-detect-compare} plot.
Surprisingly the difference between map types for the geographic
distribution was significant at the 0.05 level, even though the effect
difference is small. It also showed that the hexagon tile map display
performs marginally better than the choropleth for all trend models.

@@

\begin{itemize}
\tightlist
\item
  Add Election results data + plus other statistics,
\item
  new and better hexagon layout software
\item
  Other plot/map type comparisons, insets, cartograms, statistical plots
  along with maps, especially interactive
\item
  eye-tracking in association with lineups
\end{itemize}

### 2025-07-09 paper/review1/ReponseToReviewers2025.qmd
still to finish conclusion and working on the reason summary

@@

@@

Considering these strengths, though, as I read the manuscript I found some areas in which I would have appreciated greater clarity. Starting with some major concerns, I’ll list my comments in order of appearance:

@@

@@

@@

While there is clearly a difference between map types (Page 12, Lines 306-316), the overall level of detection rates is low. I would expect some discussion of these low results. Are these a result of the data simulation process (i.e., the “true” signal might not be different enough)? How do the results compare to the expected detection rates if selected by chance, as discussed in Section 2.2?

@@

@@

@@

I’m not sure if I understand that statement correctly. In my understanding, the interaction in Table 2 shows that the difference between choropleth and hexagon tiles differs between trend models, but it’s in favor of the hexagon models among all trends (full effect for hexagon with NW-SE trend: 1.63; full effect for hexagon all cities trend: 1.63 + 1.34 -1.16 = 1.81; full effect for hexagon tree cities trend 1.63 – 2.07 + 1.28 = 0.84).

Additionally, I can’t replicate the detection probabilities mentioned in the text, e.g., Page 13, Lines 333f.: "For the NW-SE distribution, the predicted detection rate for the hexagon tile map display increases the predicted probability of detection

to 0.63 from 0.52 for choropleths, this is almost exactly the difference seen in the table of means and is significant only at the 0.05 level."

@@

@@

@@

@@

@@

@@

*The difference is small, at least relative to the other two patterns. We expect it might be due to the large rural polygons might also play a role in obscuring large simple spatial trend, too. But we don't want to speculate.*

@@

@@

@@

@@

>>>>

*Note that we have changed the title slightly to make it more general than cancer statistics, and we have added many more references in response to the review comments. This has made the paper longer by three pages. The revision has also been converted to a quarto document, which means that all numbers are automatically calculated from the analysis code in the document.*

@@

*We are very sorry that the reviewers found the writing to be inadequate. We did not expect this! In this revision, we have tightened up the language substantially, provided additional references to support statements, and tidied the math and calculations.*

@@

Considering these strengths, though, as I read the manuscript I found some areas in which I would have appreciated greater clarity. Starting with some major concerns, I'll list my comments in order of appearance:

@@

*Added additional text.*

@@

*This has been revised.*

@@

While there is clearly a difference between map types (Page 12, Lines 306-316), the overall level of detection rates is low. I would expect some discussion of these low results. Are these a result of the data simulation process (i.e., the "true" signal might not be different enough)? How do the results compare to the expected detection rates if selected by chance, as discussed in Section 2.2?

*Sentence added to results explaining this.*

@@

*The model formulation has been revised.*

@@

*Done, in the text.*

@@

I'm not sure if I understand that statement correctly. In my understanding, the interaction in Table 2 shows that the difference between choropleth and hexagon tiles differs between trend models, but it’s in favor of the hexagon models among all trends (full effect for hexagon with NW-SE trend: 1.63; full effect for hexagon all cities trend: 1.63 + 1.34 -1.16 = 1.81; full effect for hexagon tree cities trend 1.63 – 2.07 + 1.28 = 0.84).

*Statement on interactions is removed, because the summary of the model estimates shows the different effects between the different treatments.*

Additionally, I can't replicate the detection probabilities mentioned in the text, e.g., Page 13, Lines 333f.: "For the NW-SE distribution, the predicted detection rate for the hexagon tile map display increases the predicted probability of detection to 0.63 from 0.52 for choropleths, this is almost exactly the difference seen in the table of means and is significant only at the 0.05 level."

@@

*There were errors in the hand-coded numbers. All of these are now computed using the emmeans package now.*

@@

*The explanation of data cleaning has been revised to make it clearer. There is no difference in XXX*

@@

*Language changed.*

@@

*Re-worded*

@@

*Explained.*

@@

*The difference is small, at least relative to the other two patterns. We expect it might be due to the large rural polygons might also play a role in obscuring large simple spatial trend, too. We don't want to speculate.*

@@

*A reference to Majumder, Cook and Hofmann (2025) showing that demographic background has little effect on performance of participants reading lineups. The primary variation is individual. This study was actually done in 2012 but only just published. The Vanderplas paper reports on a different type of experiment where math ability is collected as part of a larger visual testing process, and the type of plots used were ones typically seen in statistics classes. Maps are a bit different from this. Anyway, we did not collect enough data to fit demographics, because it is not the main purpose of this experiment. Majumder et al suggests that as long as a suitable range of backgrounds is included in the study results are reliable and replicable. This has been explained in the text.*

@@

*We disagree with this comment. Replicates are indicated by line segments. The primary factor, map type, is represented by a point, with paired measurements (replicates) connected by lines. This maps the primary information to position along a common axis, which is best graphical practice. Colour redundant with the facets draws attention to the secondary key element of the design, the different types of spatial patterns, and makes the plot more interesting to read.*

@@

*We have kept is separate for clarity. Speed is qute different from reason.*

@@

*Thank you for the suggestions. The Kawakami reference is out of scope. The Vanderplas et al has been included in the Introduction to point to work on how ability to read a lineup it most related to a subject's visual aptitude. Many more references to the visual inference literature have been added.*
